Introduction

### 2015-07-21 dissertation/papersincontext.tex
Excuse bad philosophy
context text

@@

@@

including the criticism forwarded by
\cite{Duan:2011:AOC:1989323.1989340} and \cite{MontoyaVCRA12},

@@

@@

>>>>

This section will put the included papers into context. It also enjoys
the benefit of hindsight, and will therefore address some of the
shortcomings that have surfaced since their publication.

@@

This work is a paper entitled ``The necessity of hypermedia RDF and an
approach to achieve it'' that was presented in the First Linked APIs
workshop at the Ninth Extended Semantic Web Conference. I was the sole
author. The reference is \cite{kjernsmo_lapis_2012}.

%TODO: not the Dagstuhl?

This paper is relevant to the topic of the dissertation for two
reasons: First, it shows how hypermedia may be used to drive a
read-write application, not just read-only, which is the focus of most
of the discussion. Second, it provides some of the basis for
\cite{ldf1}, which is important in the query planner.

The main idea in this paper is to look upon RDF as a natural language
sentence with subject, predicate and object, and that sentence should
then be sufficiently human-readable to be straightforward to
understand when a developer does ``View Source''.

Its main weakness is how it deals with authentication: The basic idea
is that when a client visits a resource, it will be told what
write operations it can perform. However, write operations will rarely
be permitted unless the client has authenticated and been authorised
to do so, and so, it assumes that credentials will be supplied with
the first read request. While the HTTP standard permits clients to
supply credentials with such requests, in practice, it rarely if ever
happens, possibly for privacy reasons. Thus, the server will have to
challenge the client first, but this should not happen on a resource
that the client is permitted to read. Thus, the proposed hypermedia
needs a property so that the client understands what to do if it wants
to do write operations.

@@

This work resulted in a paper that was nominated for Best Paper on the
Evaluation track on the International Semantic Web Conference 2013,
entitled ``Introducing Statistical Design of Experiments to SPARQL
Endpoint Evaluation'', which was co-authored with John
S. Tyssedal. The reference is \cite{kjernsmo_doe_intro}.

To understand the context of this study, I noted none of the
benchmarks that was in use for the evaluation of SPARQL engines could
be used for basic statistical hypothesis testing, and clearly,
hypothesis testing must be a fundamental requirement. Then, I noted
substantial criticism forwarded by
\cite{Duan:2011:AOC:1989323.1989340} and \cite{MontoyaVCRA12}.

However, this study merely scratches the surface. To establish a
Design of Experiment-based methodology would be a multi-year project
in its own right, clearly beyond the scope of this
dissertation. Unfortunately, this also implies that I do not have the
rigorous evaluation that I would preferred to have in this
dissertation, it is unattainable.

@@

This work resulted in a workshop paper entitled ``Pushing complexity
down the stack'', presented in the Developers Workshop at the
International Semantic Web Conference, by the lead author Gregory Todd
Williams, we were the two authors. The reference is
\cite{williamspushing}.

% TODO: detailed contribution?

This work was the outcome of several hackathons that I had organised
in the Perl community. It is not very well known that Gregory Todd
Williams' SPARQL implementation written in Perl is one of the SPARQL
1.1 reference implementations, and is known to be fully compliant with
the specification. Through our work, and the availability of
traits-based programming in Perl, see \cite{traits}, we had extensive
discussions on how to leave optimisations to lower levels, e.g. the
underlying database engine, when they had the potential to provide a
better plan or cost estimate. While it is certainly possible to do
this without this new programming paradigm, traits made the
methodology much more comprehensible, and so helped the subsequent
work substantially.

This work also resulted in the ``Attean''
framework\footnote{\url{https://metacpan.org/release/Attean}}, also
mainly authored by Gregory Todd Williams with some smaller
contributions from me. It is also available in the Linux distribution
Debian (and derivatives, such as Ubuntu) as \texttt{libattean-perl}.

@@

\subsection{Philosophy of Science}\label{sec:conphil}
